---
title: "Cycling survey results"
author: "Steve"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(fs)
library(readxl)
library(scales)

plot_folder <- "../Bendigo survey/plots"
if (!dir.exists(plot_folder)) dir_create(plot_folder)

```

```{r load_data, message=FALSE}
surveyFile <- "../Bendigo survey/responses-8gi6nyx69dt6-2025-07-10T04_40_41.798Z.xlsx"

sheet_names <- excel_sheets(surveyFile)

spot_sheet_names <- c("Most stressful intersection(s) ",
                      "My favorite spot(s) along the r",
                      "My least favorite spot(s) along")

route_sheet_names <- sheet_names[!sheet_names %in% c("Respondents", 
                                                     spot_sheet_names)]

respondents <- read_excel(surveyFile, sheet = "Respondents")

routes <- lapply(route_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$destination <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()

spots <-  lapply(spot_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$type <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()


```

# Introduction
TO DO

# Sociodemographic characteristics
This section provides general information about the survey participants.

## Ridership

```{r}
q <- respondents %>% select(starts_with("Are you a bicycle rider")) %>% names()

categorise <- function(x) {
  case_when(
    str_detect(x, "once a week")  ~ "Weekly",
    str_detect(x, "once a month") ~ "Monthly",
    str_detect(x, "occasionally") ~ "Occasionally",
    str_detect(x, "not cycled")   ~ "Not cycled",
    is.na(x)                      ~ "Did not answer",
    TRUE                          ~ NA_character_
  )
}
data <- respondents %>%
  select(all_of(q)) %>%
  mutate(category = categorise(.data[[q]])) %>%
  mutate(category = factor(category,
                           levels = c("Weekly", "Monthly", "Occasionally",
                                      "Not cycled", "Did not answer")))
responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  arrange(factor(category,
                 levels = c("Weekly","Monthly","Occasionally",
                            "Not cycled","Did not answer"))) %>%
  mutate(full = paste0(category, ": '", response, "'")) %>%
  pull(full) %>%
  paste(collapse = "\n")

valid_responses <- respondents %>%
  mutate(category = categorise(.data[[q]])) %>%
  filter(category %in% c("Weekly", "Monthly", "Occasionally"))
```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Category",
    y       = "Number of respondents",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.0 ridership.png"), plot,
       width = 18, height = 12, units = "cm")

```

Out of `r nrow(respondents)` survey participants, `r nrow(respondents) - nrow(valid_responses)` answered that they had not cycled in the past year, or did not answer. These were excluded from further survey questions.

The results presented below for the remainder of the sociodemographic questions exclude participants who did not answer the relevant question.


## Gender

```{r}
q <- respondents %>% select(starts_with("1.1")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(gender = factor(.data[[q]],
                         levels = c("Woman/Female", "Man/Male", 
                                    "Other", "Prefer not to say")))

```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = gender)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Gender",
    y       = "Number of participants"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.1 gender.png"), plot,
       width = 18, height = 12, units = "cm")

```

## Age

```{r, warning=FALSE}
q <- respondents %>% select(starts_with("1.2")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  
  # anomalous answers - assign midpoint
  mutate(age_num = case_when(
    str_detect(.data[[q]], "Mid 40s") ~ 45,
    str_detect(.data[[q]], "50’s") ~ 55,
    TRUE ~ as.numeric(.data[[q]])
  )) %>%
  # ambiguous answers/errors (60+, 565) - exclude
  filter(!is.na(age_num) & age_num <= 100) %>%
  
  mutate(age_group = cut(
  age_num,
      breaks = c(17, 20, 30, 40, 50, 60, 70, 80, Inf),
      labels = c("18-20", "21-30", "31-40", "41-50",
                 "51-60", "61-70", "71-80", "80+"),
      right = TRUE,   # intervals are closed on the right
      include.lowest = TRUE
    ))
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = age_group)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Age group",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.2 age.png"), plot,
       width = 18, height = 12, units = "cm")

```

## Living arrangements

```{r}
q <- respondents %>% select(starts_with("1.3")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(living_arrangement = .data[[q]] %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev() %>%   # reverse so largest is on top in horizontal bar chart
           fct_relevel("Prefer not to say", after = 0)# push to first (bottom in horizontal bars)
  )
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = living_arrangement)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Living arrangement",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.3 living.png"), plot,
       width = 18, height = 9, units = "cm")

```

## Cars

```{r}
q <- respondents %>% select(starts_with("1.4")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_cars = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = num_cars)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of cars",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.4 cars.png"), plot,
       width = 18, height = 12, units = "cm")

```

## Bicycles

```{r}
q <- respondents %>% select(starts_with("1.5")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_bikes = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = num_bikes)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of bicycles",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.5 bikes.png"), plot,
       width = 18, height = 12, units = "cm")

```

## Educational qualifications

```{r}
q <- respondents %>% select(starts_with("1.6")) %>% names()
col_names <- names(respondents)
next_col <- col_names[which(col_names == q) + 1]

categorise <- function(x, y) {
  case_when(
    str_detect(x, "Year 10")         ~ "Year 10 or less",
    x == "Other" & (str_detect(y, "11") | str_detect(y, "12")) ~ "Year 11 or 12",
    str_detect(x, "Diploma") ~ "Diploma/certificate",
    str_detect(x, "Undergraduate")   ~ "Undergraduate",
    str_detect(x, "Postgraduate")    ~ "Postgraduate",
    x == "Other"                     ~ "Other",
    TRUE                             ~ NA_character_
  )
}

data <- valid_responses %>%
  select(all_of(c(q, next_col))) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(qualif = categorise(.data[[q]], .data[[next_col]])) %>%
  mutate(qualif = factor(qualif,
                           levels = c("Year 10 or less", "Year 11 or 12", "Diploma/certificate",
                                      "Undergraduate", "Postgraduate", "Other")))

responses <- data %>%
  distinct() %>%
  rename(response = 1) %>%arrange(qualif) %>%
  filter(str_detect(qualif, "Diploma") | str_detect(qualif, "Undergraduate") | 
           str_detect(qualif, "Postgraduate"))

caption_text <- responses %>%
  arrange(factor(qualif,
                 levels = c("Diploma/certificate", "Undergraduate", "Postgraduate"))) %>%
  mutate(full = paste0(qualif, ": '", response, "'")) %>%
  pull(full) %>%
  paste(collapse = "\n")

```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = qualif)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Educational qualification",
    y = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.6 qualification.png"), plot,
       width = 18, height = 12, units = "cm")

```

## Occupation

```{r}
q <- respondents %>% select(starts_with("1.7")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]]) & !(.data[[q]] == "-")) %>%
  mutate(occup = tolower(.data[[q]])) %>%
  mutate(occup = case_when(
    str_detect(occup, "retired") & !str_detect(occup, "semi") ~ "retired",
    occup == "admin" ~ "administration",
    occup == "physio" ~ "physiotherapist",
    TRUE             ~ occup
  )) %>%
  mutate(occup = str_to_sentence(occup)) %>%
  mutate(occup = case_when(
    occup == "Ceo"                           ~ "CEO",
    occup == "Council officer (not cogb)"    ~ "Council officer (not CoGB)",
    occup == "Early support coordinator - transport accident commission" ~
      "Early support coordinator - Transport Accident Commission",
    occup == "Gis officer"                   ~ "GIS officer",
    occup == "It administrator/ gis analyst" ~ "IT administrator / GIS analyst",
    occup == "Gp"                            ~ "GP",
    occup == "Mri applications specialist"   ~ "MRI applications specialilst",
    occup == "Public servant dtp"            ~ "Public servant DTP",
    TRUE                                     ~ occup
  )) %>%
  group_by(occup) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  arrange(desc(n), occup)

data_plot <- data %>%
  filter(n >= 3) %>%
  mutate(occup = occup %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev())  # reverse so largest is on top in horizontal bar chart

others_text <- data %>%
  filter(n < 3) %>%
  mutate(text = paste0(occup, " (", n, ")")) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = ", ")


```

**`r q`**

```{r}
plot <- ggplot(data_plot, aes(x = occup)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Occupation",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.7 occupation.png"), plot,
       width = 18, height = 9, units = "cm")

```

Other occupations not included in the plot above: 
`r others_text`.

## Income

```{r}
q <- respondents %>% select(starts_with("1.8")) %>% names()

format_income <- function(x) {
  str_replace_all(x, 
                  "(\\d+)",  # match any sequence of digits
                  function(y) paste0("$", comma(as.numeric(y))))
}

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(income_label = format_income(.data[[q]])) %>%
  mutate(income_order = case_when(
    str_detect(income_label, "Below") ~ 0,
    str_detect(income_label, "Above") ~ 200000,  # arbitrarily above highest range
    str_detect(income_label, "I prefer") ~ Inf,
    TRUE ~ as.numeric(str_extract(income_label, "\\d+"))
  )) %>%
  mutate(income_label = fct_reorder(income_label, income_order) %>%
           fct_rev())
  
```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = income_label)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Household income",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.8 income.png"), plot,
       width = 18, height = 9, units = "cm")

```

# Cycling experiences

This section provides information about routes drawn by participants.

## Route purpose
Participants were asked to describe the purpose of their route.

```{r}

## TO do - graph of the preferred route categories

```

## Reasons for cycling


## Reasons for selecting route



