---
title: "Cycling survey results"
author: "Steve"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(fs)
library(readxl)
library(scales)

plot_folder <- "../Bendigo survey/plots"
if (!dir.exists(plot_folder)) dir_create(plot_folder)

```

```{r load_data, message=FALSE}
surveyFile <- "../Bendigo survey/responses-8gi6nyx69dt6-2025-07-10T04_40_41.798Z.xlsx"

sheet_names <- excel_sheets(surveyFile)

spot_sheet_names <- c("Most stressful intersection(s) ",
                      "My favorite spot(s) along the r",
                      "My least favorite spot(s) along")

route_sheet_names <- sheet_names[!sheet_names %in% c("Respondents", 
                                                     spot_sheet_names)]

respondents <- read_excel(surveyFile, sheet = "Respondents")

routes <- lapply(route_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$destination <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()

spots <-  lapply(spot_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$type <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()

```

# Introduction

*[Include some details about how the survey was conducted, eg who it was conducted for (presumably City of Greater Bendigo), who conducted it, the dates it was conducted, and how participants were invited to participate.]*

There were `r nrow(respondents)` survey responses, though some participants did not answer all questions.  

The survey is divided into 3 parts:
- sociodemographic characteristics
- cycling experiences
- cycling preferences
 
# 1. Sociodemographic characteristics
This section provides general information about the survey participants.

### Ridership

```{r}

categorise <- function(x) {
  case_when(
    str_detect(x, "once a week")  ~ "Weekly",
    str_detect(x, "once a month") ~ "Monthly",
    str_detect(x, "occasionally") ~ "Occasionally",
    str_detect(x, "not cycled")   ~ "Not cycled",
    is.na(x)                      ~ "Did not answer",
    TRUE                          ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("Are you a bicycle rider")) %>% names()

data <- respondents %>%
  select(all_of(q)) %>%
  mutate(category = categorise(.data[[q]]),
         category = factor(category,
                           levels = c("Weekly", "Monthly", "Occasionally",
                                      "Not cycled", "Did not answer")))
responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  filter(category != "Did not answer") %>% 
  arrange(factor(category,
                 levels = c("Weekly", "Monthly", "Occasionally", "Not cycled"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

valid_responses <- respondents %>%
  mutate(category = categorise(.data[[q]])) %>%
  filter(category %in% c("Weekly", "Monthly", "Occasionally"))
```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Category",
    y       = "Number of respondents",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.0 ridership.png"), plot,
       width = 18, height = 9, units = "cm")

```

Out of `r nrow(respondents)` survey participants, `r nrow(respondents) - nrow(valid_responses)` answered that they had not cycled in the past year, or did not answer. These were excluded from further survey questions.

The results presented below for the remainder of the sociodemographic questions exclude participants who did not answer the relevant question.


### Gender

```{r}
q <- respondents %>% select(starts_with("1.1")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(gender = factor(.data[[q]],
                         levels = c("Woman/Female", "Man/Male", 
                                    "Other", "Prefer not to say")))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = gender)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Gender",
    y       = "Number of participants"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.1 gender.png"), plot,
       width = 18, height = 9, units = "cm")

```

There were slightly more participants describing themselves as male (`r round(nrow(data %>% filter(gender == "Man/Male")) / nrow(data) * 100)`%) than as women or female (`r round(nrow(data %>% filter(gender == "Woman/Female")) / nrow(data) * 100)`%).

### Age

```{r, warning=FALSE}
q <- respondents %>% select(starts_with("1.2")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  
  # anomalous answers - assign midpoint
  mutate(age_num = case_when(
    str_detect(.data[[q]], "Mid 40s") ~ 45,
    str_detect(.data[[q]], "50’s") ~ 55,
    TRUE ~ as.numeric(.data[[q]])
  )) %>%
  # ambiguous answers/errors (60+, 565) - exclude
  filter(!is.na(age_num) & age_num <= 100) %>%
  
  mutate(age_group = cut(
  age_num,
      breaks = c(17, 20, 30, 40, 50, 60, 70, 80, Inf),
      labels = c("18-20", "21-30", "31-40", "41-50",
                 "51-60", "61-70", "71-80", "80+"),
      right = TRUE,   # intervals are closed on the right
      include.lowest = TRUE
    ))
    
```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = age_group)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Age group",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.2 age.png"), plot,
       width = 18, height = 9, units = "cm")

```

Most participants (`r round(nrow(data %>% filter(age_num > 30)) / nrow(data) * 100)`%) were aged over 30, including a large group aged over 60 (`r round(nrow(data %>% filter(age_num > 60)) / nrow(data) * 100)`%). 

### Living arrangements

```{r}
q <- respondents %>% select(starts_with("1.3")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(living_arrangement = .data[[q]] %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev() %>%   # reverse so largest is on top in horizontal bar chart
           fct_relevel("Prefer not to say", after = 0)# push to first (bottom in horizontal bars)
  )
    
```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = living_arrangement)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Living arrangement",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.3 living.png"), plot,
       width = 18, height = 9, units = "cm")

```

Most participants (`r round(nrow(data %>% filter(living_arrangement %in% c("Couple with children", "Couple, no children"))) / nrow(data) * 100)`%) were in couples, and almost half (`r round(nrow(data %>% filter(living_arrangement %in% c("Couple with children", "Lone parent with child/children"))) / nrow(data) * 100)`%) were in families with children.

### Cars

```{r}
q <- respondents %>% select(starts_with("1.4")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_cars = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = num_cars)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of cars",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.4 cars.png"), plot,
       width = 18, height = 9, units = "cm")

```

Around `r round(nrow(data %>% filter(num_cars == 2)) / nrow(data) * 100)`% of households had 2 cars, with only around `r round(nrow(data %>% mutate(num_cars = as.numeric(as.character(num_cars))) %>% filter(num_cars > 2)) / nrow(data) * 100)`% having more than 2 cars.


### Bicycles

```{r}
q <- respondents %>% select(starts_with("1.5")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_bikes = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = num_bikes)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of bicycles",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.5 bikes.png"), plot,
       width = 18, height = 9, units = "cm")

```

Around `r round(nrow(data %>% filter(num_bikes == 2)) / nrow(data) * 100)`% of households had 2 bicycles, with around `r round(nrow(data %>% mutate(num_bikes = as.numeric(as.character(num_bikes))) %>% filter(num_bikes > 2)) / nrow(data) * 100)`% having more than 2 bicycles.


### Educational qualifications

```{r}
categorise <- function(x, y) {
  case_when(
    str_detect(x, "Year 10")         ~ "Year 10 or less",
    x == "Other" & (str_detect(y, "11") | str_detect(y, "12")) ~ "Year 11 or 12",
    str_detect(x, "Diploma") ~ "Diploma/certificate",
    str_detect(x, "Undergraduate")   ~ "Undergraduate",
    str_detect(x, "Postgraduate")    ~ "Postgraduate",
    x == "Other"                     ~ "Other",
    TRUE                             ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("1.6")) %>% names()
col_names <- names(respondents)
next_col <- col_names[which(col_names == q) + 1]

data <- valid_responses %>%
  select(all_of(c(q, next_col))) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(qualif = categorise(.data[[q]], .data[[next_col]])) %>%
  mutate(qualif = factor(qualif,
                           levels = c("Year 10 or less", "Year 11 or 12", "Diploma/certificate",
                                      "Undergraduate", "Postgraduate", "Other")))

responses <- data %>%
  distinct() %>%
  rename(response = 1) %>%
  arrange(qualif) %>%
  filter(str_detect(qualif, "Diploma") | str_detect(qualif, "Undergraduate") | 
           str_detect(qualif, "Postgraduate"))

caption_text <- responses %>%
  arrange(factor(qualif,
                 levels = c("Diploma/certificate", "Undergraduate", "Postgraduate"))) %>%
  mutate(full = paste0(qualif, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = qualif)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Educational qualification",
    y = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.6 qualification.png"), plot,
       width = 18, height = 9, units = "cm")

```

Most participants (`r round(nrow(data %>% filter(qualif %in% c("Diploma/certificate", "Undergraduate", "Postgraduate"))) / nrow(data) * 100)`%) reported a tertiary qualification of some kind.


### Occupation

```{r}
q <- respondents %>% select(starts_with("1.7")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]]) & !(.data[[q]] == "-")) %>%
  mutate(occup = tolower(.data[[q]])) %>%
  mutate(occup = case_when(
    str_detect(occup, "retired") & !str_detect(occup, "semi") ~ "retired",
    occup == "admin" ~ "administration",
    occup == "physio" ~ "physiotherapist",
    TRUE             ~ occup
  )) %>%
  mutate(occup = str_to_sentence(occup)) %>%
  mutate(occup = case_when(
    occup == "Ceo"                           ~ "CEO",
    occup == "Council officer (not cogb)"    ~ "Council officer (not CoGB)",
    occup == "Early support coordinator - transport accident commission" ~
      "Early support coordinator - Transport Accident Commission",
    occup == "Gis officer"                   ~ "GIS officer",
    occup == "It administrator/ gis analyst" ~ "IT administrator / GIS analyst",
    occup == "Gp"                            ~ "GP",
    occup == "Mri applications specialist"   ~ "MRI applications specialilst",
    occup == "Public servant dtp"            ~ "Public servant DTP",
    TRUE                                     ~ occup
  )) %>%
  group_by(occup) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  arrange(desc(n), occup)

data_plot <- data %>%
  filter(n >= 3) %>%
  mutate(occup = occup %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev())  # reverse so largest is on top in horizontal bar chart

others_text <- data %>%
  filter(n < 3) %>%
  mutate(text = paste0(occup, " (", n, ")")) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = ", ")


```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data_plot, aes(x = occup)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Occupation",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.7 occupation.png"), plot,
       width = 18, height = 9, units = "cm")

```

Other occupations not included in the plot above: 
`r others_text`.

A wide variety of occupations were reported, with 'Retired' (`r round(nrow(data %>% filter(occup == "Retired")) / nrow(data) * 100)`%) the largest single reported occupation, consistently with the large number of participants aged over 60.

### Income

```{r}
q <- respondents %>% select(starts_with("1.8")) %>% names()

format_income <- function(x) {
  str_replace_all(x, 
                  "(\\d+)",  # match any sequence of digits
                  function(y) paste0("$", comma(as.numeric(y))))
}

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(income_label = format_income(.data[[q]])) %>%
  mutate(income_order = case_when(
    str_detect(income_label, "Below") ~ 0,
    str_detect(income_label, "Above") ~ 200000,  # arbitrarily above highest range
    str_detect(income_label, "I prefer") ~ Inf,
    TRUE ~ as.numeric(str_extract(income_label, "\\d+"))
  )) %>%
  mutate(income_label = fct_reorder(income_label, income_order) %>%
           fct_rev())
  
```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = income_label)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Household income",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.8 income.png"), plot,
       width = 18, height = 9, units = "cm")

```

The most common income category was \$45,001 to \$135,000 (`r round(nrow(data %>% filter(income_label == "$45,001 - $135,000")) / nrow(data) * 100)`% of participants).


# 2. Cycling experiences

This section provides information about routes drawn by participants.  A total of `r nrow(routes)` route responses were received from by participants, though 13 were of either zero length or under 50m, leaving a total of `r nrow(routes) - 13` valid routes. 

### Route purpose
Participants were asked to describe the purpose of their route.

```{r fig.width = 8, fig.height = 4}
clean_destination <- function(x) {
  x <- case_when(
    x == "To public transport stop-statio" ~ "To public transport stop/station",
    TRUE ~ x
  )
  x <- gsub("-", "/", x)
  x <- gsub(", ", "/", x)
  x <- factor(
    x,
    levels = c(
      "To work/commute", "To college/university/classes",
      "To shops/tasks/errands", "To visit family/friends",
      "To public transport stop/station", "To other destinations"
    )
  )
  return(x)
}

data <- routes %>% 
  dplyr::select(`Respondent ID`, destination, `Please tell us the type of destination:`) %>%
  mutate(destination = clean_destination(destination) %>%
           fct_rev())

others_text <- data %>%
  filter(!is.na(`Please tell us the type of destination:`)) %>%
  group_by(`Please tell us the type of destination:`) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(`Please tell us the type of destination:`, " (", n, ")"),
                       `Please tell us the type of destination:`)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")


plot <- ggplot(data, aes(x = destination, fill = destination)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Trip purpose",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none")

plot
ggsave(paste0(plot_folder, "/2.0 purpose.png"), plot,
       width = 18, height = 9, units = "cm")

```

Destinations stated by participants under 'Other':<br>
`r others_text`

The largest category of trip purpose was 'other' (`r round(nrow(data %>% filter(destination == "To other destinations")) / nrow(data) * 100)`%), followed by 'work/commute' (`r round(nrow(data %>% filter(destination == "To work/commute")) / nrow(data) * 100)`%). While a wide variety of 'other' destinations were stated, many of them relate to recreational trips.


### Reasons for cycling

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "exercise")   ~ "Exercise",
    str_detect(x, "enjoy")      ~ "Enjoyment",
    str_detect(x, "available")  ~ "Availability",
    str_detect(x, "convenient") ~ "Convenience",
    str_detect(x, "peers")      ~ "Peers",
    str_detect(x, "Other")      ~ "Other",
    TRUE                        ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "1.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

responses <- data.frame(
  response = str_replace(q_cols, "^.*?:\\s*", ""),  # everything after colon
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Exercise", "Enjoyment", "Availability",
                                            "Convenience", "Peers", "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Exercise", "Enjoyment", "Availability",
                                            "Convenience", "Peers", "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 5}
plot <- ggplot(data, 
               aes(x = category, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Reason for cycling",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))


plot
ggsave(paste0(plot_folder, "/2.1.1 reason.png"), plot,
       width = 18, height = 12, units = "cm")


```

Reasons stated by participants under 'Other':<br>
`r others_text`

Exercise (`r round(nrow(data %>% filter(category == "Exercise")) / nrow(data) * 100)`%) and enjoyment (`r round(nrow(data %>% filter(category == "Enjoyment")) / nrow(data) * 100)`%) were the most commonly-stated reasons, followed by convenience (`r round(nrow(data %>% filter(category == "Convenience")) / nrow(data) * 100)`%). Few participants (`r round(nrow(data %>% filter(category == "Availability")) / nrow(data) * 100)`%) were driven by the necessity of bicycle being the only available method.

### Reasons for selecting route

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "safe")          ~ "Safe",
    str_detect(x, "greenery")      ~ "Green",
    str_detect(x, "flat")          ~ "Flat",
    str_detect(x, "traffic")       ~ "Traffic",
    str_detect(x, "intersections") ~ "Intersections",
    str_detect(x, "speed")         ~ "Speed",
    str_detect(x, "connectivity")  ~ "Connectivity",
    str_detect(x, "shortest")      ~ "Short",
    str_detect(x, "cyclists")      ~ "Cyclists",
    str_detect(x, "designated")    ~ "Designated",
    str_detect(x, "Other")         ~ "Other",
    TRUE                           ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "2.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], ".*\\)")  # part of col name before )

responses <- data.frame(
  response = str_replace(q_cols, "^.*?\\):\\s*", ""),  # everything after ):
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?\\):\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Safe", "Green", "Flat", "Traffic", 
                            "Intersections", "Speed", "Connectivity",  
                            "Short","Cyclists", "Designated", "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Safe", "Green", "Flat", "Traffic", 
                            "Intersections", "Speed", "Connectivity",  
                            "Short","Cyclists", "Designated", "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.1.`r q`**

```{r  fig.width = 8, fig.height = 6}
plot <- ggplot(data,
               aes(x = category, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Reason for choosing route",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") 


plot
ggsave(paste0(plot_folder, "/2.1.2 route choice.png"), plot,
       width = 18, height = 15, units = "cm")

```

Reasons stated by participants under 'Other':<br>
`r others_text`.

Safety (`r round(nrow(data %>% filter(category == "Safe")) / nrow(data) * 100)`%), low traffic (`r round(nrow(data %>% filter(category == "Traffic")) / nrow(data) * 100)`%), shortness of route (`r round(nrow(data %>% filter(category == "Short")) / nrow(data) * 100)`%) and availability of designated bike paths (`r round(nrow(data %>% filter(category == "Designated")) / nrow(data) * 100)`%) were the highest-ranked reasons for selecting routes.

### Maximum distance

```{r}
q <- routes %>% select(starts_with("3.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(distance = str_replace(.data[[q]], "(km)", "\\1\n"),  # insert line break after 'km'
         distance = str_replace(distance, "approximately ", "~"),
         distance = str_replace(distance, "more than ", ">"),
         distance = factor(distance,
                         levels = c("Up to 5 km\n (~15 minutes)",
                                    "Up to 10 km\n (~30 minutes)",
                                    "Up to 20 km\n (~1 hour)",
                                    "Up to 30 km\n (~1.5 hour)",
                                    "More than 30 km\n (>1.5 hour)")),
         
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, 
               aes(x = distance, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Distance willing to travel",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/2.1.3 distance.png"), plot,
       width = 18, height = 9, units = "cm")

```

'Up to 10 km (~30 minutes)' was the most common distance (`r round(nrow(data %>% filter(distance == "Up to 10 km\n (~30 minutes)")) / nrow(data) * 100)`%) participants would be willing to travel for their stated purpose. Participants were not  willing to travel for more than 30 km (> 1.5 hour) for work/commuting or education purposes, with most trips where that distance was acceptable being 'other' (typically recreational) trips.

### Unfavourable route aspects

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "unsafe")        ~ "Safety",
    str_detect(x, "separation")    ~ "Separation",
    str_detect(x, "greenery")      ~ "Greenery",
    str_detect(x, "traffic")       ~ "Traffic",
    str_detect(x, "intersections") ~ "Intersections",
    str_detect(x, "speed")         ~ "Speed",
    str_detect(x, "connectivity")  ~ "Connectivity",
    str_detect(x, "surface")       ~ "Surface",
    str_detect(x, "steep")         ~ "Steepness",
    str_detect(x, "Other")         ~ "Other",
    TRUE                           ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "4.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

responses <- data.frame(
  response = str_replace(q_cols, "^.*?:\\s*", ""),  # everything after colon
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Safety", "Separation", "Greenery",
                                           "Traffic", "Intersections", "Speed",
                                           "Connectivity", "Surface", "Steepness",
                                           "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Safety", "Separation", "Greenery",
                            "Traffic", "Intersections", "Speed",
                            "Connectivity", "Surface", "Steepness",
                            "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 5}
plot <- ggplot(data, 
               aes(x = category, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Unfavourable route aspects",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom")

plot
ggsave(paste0(plot_folder, "/2.1.4 unfavourable.png"), plot,
       width = 18, height = 15, units = "cm")

```

Unfavourable route aspects stated by participants under 'Other':<br>
`r others_text`

Lack of separation from traffic (`r round(nrow(data %>% filter(category == "Separation")) / nrow(data) * 100)`%), poor connectivity (`r round(nrow(data %>% filter(category == "Connectivity")) / nrow(data) * 100)`%), heavy traffic (`r round(nrow(data %>% filter(category == "Traffic")) / nrow(data) * 100)`%), and poor safety (`r round(nrow(data %>% filter(category == "Safety")) / nrow(data) * 100)`%) were the most frequently-identified unsatisfactory route aspects. 

### Frequency of cycling route

```{r}
q <- routes %>% select(starts_with("5.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(frequency = str_replace(.data[[q]], "-", "-\n"),  # insert line break after first dash
         frequency = factor(frequency,
                         levels = c("Frequently-\n 5 days or more per week",
                                    "Regularly-\n at least 4 days per week",
                                    "Often-\n 1-3 days per week",
                                    "From time to time-\n 1-4 days per month")),
         
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 5}
plot <- ggplot(data, 
               aes(x = frequency, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Frequency of cycling route",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        legend.position = "bottom")

plot
ggsave(paste0(plot_folder, "/2.1.5 frequency.png"), plot,
       width = 18, height = 12, units = "cm")

```

The most common answers were 1-4 days per month (`r round(nrow(data %>% filter(frequency == "From time to time-\n 1-4 days per month")) / nrow(data) * 100)`%), followed by 1-3 days per week (`r round(nrow(data %>% filter(frequency == "Often-\n 1-3 days per week")) / nrow(data) * 100)`%).

### Satisfaction level

```{r}
q <- routes %>% select(starts_with("6.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(satisf = str_replace(.data[[q]], "-", "-\n"),  # insert line break after first dash
         satisf = factor(satisf,
                         levels = c("Very satisfied",
                                    "Satisfied",
                                    "Neutral",
                                    "Dissatisfied",
                                    "Very dissatisfied")),
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, 
               aes(x = satisf, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Satisfaction with route",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/2.1.6 satisfaction.png"), plot,
       width = 18, height = 9, units = "cm")

```

Around `r round(nrow(data %>% filter(satisf %in% c("Satisfied", "Very satisfied"))) / nrow(data) * 100)`% of participants described themselves as 'Satisfied' or 'Very satisfied'.

### Accompanying people

```{r}
col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "7.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "people", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    people = ifelse(people == "Other family members",
                    "Other family\n members", people),
    people = factor(people, levels = c("I cycle alone", "My friend(s)", "My partner",
                                        "My kids", "Other family\n members"))
  )


```

**2.1.`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, 
               aes(x = people, fill = destination)) +
  geom_bar(position = "stack") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Accompanying people",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/2.1.7 accompanying.png"), plot,
       width = 18, height = 9, units = "cm")


```

The most common answer was 'I cycle alone' (`r round(nrow(data %>% filter(people == "I cycle alone")) / nrow(data) * 100)`%), and this answer was given for `r round(nrow(data %>% filter(people == "I cycle alone" & destination == "To work/commute")) / nrow(data %>% filter(destination == "To work/commute")) * 100)`% of work/commute trips.

```{r}
spot_numbers <- spots %>%
  mutate(category = case_when(
    str_detect(type, "My favorite")    ~ "Favourite",
    str_detect(type, "least favorite") ~ "Least favourite",
    str_detect(type, "stressful")       ~ "Stressful",
    TRUE                                ~ NA_character_
  )) %>%
  group_by(category) %>%
  summarise(n = n()) %>%
  ungroup()

fav_no <- spot_numbers[spot_numbers$category == "Favourite", "n"]

fav_no <- spot_numbers %>% filter(category == "Favourite") %>% pull(n)
```


### Favourite spot

**2.2 My favorite spot(s) along the route**

A total of `r spot_numbers %>% filter(category == "Favourite") %>% pull(n)` favourite spots were marked by participants.

### Stressful intersections

**2.3 Most stressful instersection(s)**

A total of `r spot_numbers %>% filter(category == "Stressful") %>% pull(n)` stressful intersections were marked by participants.

### Least favourite spot

**2.4 My least favorite spot(s) along the route**

A total of `r spot_numbers %>% filter(category == "Least favourite") %>% pull(n)` least favourite spots were marked by participants.

### Barriers and challenges

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "safe cycling")        ~ "Unsafe infrastructure",
    str_detect(x, "Heavy traffic")       ~ "Heavy traffic",
    str_detect(x, "Fast traffic")        ~ "Fast traffic",
    str_detect(x, "bike parking")        ~ "Lack of bike parking",
    str_detect(x, "personal safety")     ~ "Personal safety",
    str_detect(x, "lighting")            ~ "Lighting",
    str_detect(x, "Poor road")           ~ "Road conditions",
    str_detect(x, "separated cycling")   ~ "Lack of separation",
    str_detect(x, "Hilly terrain")       ~ "Hilliness",
    str_detect(x, "changing facilities") ~ "End-of-trip facilities",
    str_detect(x, "Weather")             ~ "Weather",
    str_detect(x, "maintenance")         ~ "Lack of equipment",
    str_detect(x, "awareness")           ~ "Lack of awareness",
    str_detect(x, "health")              ~ "Health",
    str_detect(x, "Convenience")         ~ "Convenience",
    str_detect(x, "signs")               ~ "Lack of signage",
    str_detect(x, "greenery")            ~ "Lack of greenery",
    str_detect(x, "Other")               ~ "Other",
    TRUE                                 ~ NA_character_
  )
}

col_names <- names(respondents)
q_cols <- col_names[startsWith(col_names, "2.5")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column

q <- stringr::str_extract(q_cols[1], ".*\\)")  # part of col name before )

responses <- data.frame(
  response = str_replace(q_cols, "^.*?\\):\\s*", ""),  # everything after ):
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- valid_responses %>% 
  select(all_of(q_cols)) %>%
  rename_with(~ str_replace(.x, "^.*?\\):\\s*", "")) %>% 
  pivot_longer(everything(), names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    category = categorise(reason),
    category = factor(category, levels = c("Unsafe infrastructure", "Heavy traffic",
                                           "Fast traffic", "Lack of bike parking",
                                           "Personal safety", "Lighting",
                                           "Road conditions", "Lack of separation",
                                           "Hilliness", "End-of-trip facilities",
                                           "Weather", "Lack of equipment",
                                           "Lack of awareness", "Health",
                                           "Convenience", "Lack of signage",
                                           "Lack of greenery", "Other")),
    category = fct_rev(category)
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Unsafe infrastructure", "Heavy traffic",
                            "Fast traffic", "Lack of bike parking",
                            "Personal safety", "Lighting",
                            "Road conditions", "Lack of separation",
                            "Hilliness", "End-of-trip facilities",
                            "Weather", "Lack of equipment",
                            "Lack of awareness", "Health",
                            "Convenience", "Lack of signage",
                            "Lack of greenery", "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- respondents %>%
  select(any_of(other_col)) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.`r q`**

```{r fig.width = 8, fig.height = 8}
plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Barrier or challenge",
    y = "Number of participants",
    caption = caption_text
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        plot.caption.position = "plot")

plot
ggsave(paste0(plot_folder, "/2.5 challenges.png"), plot,
       width = 18, height = 18, units = "cm")

```

Barriers and challenges stated by participants under 'Other':<br> 
`r others_text`.

'Lack of separation' (`r round(nrow(data %>% filter(category == "Lack of separation")) / nrow(data) * 100)`%) and 'unsafe infrastructure' (`r round(nrow(data %>% filter(category == "Unsafe infrastructure")) / nrow(data) * 100)`%) were the most frequently-stated barriers and challenges, emphasising the high importance of safety considerations to participants.

# 3. Cycling preferences

This section provides information about preferences stated by participants.

### Bicycle type

```{r}
q <- respondents %>% select(starts_with("3.1 ")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(bicycle = factor(.data[[q]],
                         levels = c("Classic bicycle", "E-bike", 
                                    "Both")))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = bicycle)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Type of bicycle",
    y       = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.1 bicycle.png"), plot,
       width = 18, height = 9, units = "cm")

```

'Classic bicycle' (`r round(nrow(data %>% filter(bicycle == "Classic bicycle")) / nrow(data) * 100)`%) was the most common answer.

### Years riding

```{r}
q <- respondents %>% select(starts_with("3.2")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(years = factor(.data[[q]],
                         levels = c("Less than 1 year", "1-2 years",
                                    "3-5 years", "6-10 years",
                                    "More than 10 years")))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = years)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Years riding",
    y       = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.2 years.png"), plot,
       width = 18, height = 9, units = "cm")

```

Participants were typically very experienced riders, with around `r round(nrow(data %>% filter(years == "More than 10 years")) / nrow(data) * 100)`% having ridden for more than 10 years.

### Satisfaction with infrastructure

```{r}
q <- respondents %>% select(starts_with("3.3")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(satisf = factor(.data[[q]],
                         levels = c("Very satisfied", "Satisfied",
                                    "Neutral", "Dissatisfied",
                                    "Very dissatisfied")))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = satisf)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Satisfaction with infrastructure",
    y       = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.3 satisfaction.png"), plot,
       width = 18, height = 9, units = "cm")

```

Responses to this question were split, with around `r round(nrow(data %>% filter(satisf %in% c("Satisfied", "Very satisfied"))) / nrow(data) * 100)`% expressing themselves as 'Satisfied' or 'Very satisfied', while around `r round(nrow(data %>% filter(satisf %in% c("Dissatisfied", "Very dissatisfied"))) / nrow(data) * 100)`% were 'Dissatisfied' or 'Very dissatisfied'.

### Cycle friendliness

```{r}
q <- respondents %>% select(starts_with("3.4")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(friend = factor(.data[[q]],
                         levels = c("Very cyclist-friendly", "Somewhat friendly to cyclists",
                                    "Neutral", "Somewhat unfriendly to cyclists",
                                    "Very unfriendly to cyclists")),
         friend = fct_rev(friend))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = friend)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Cycle friendliness",
    y       = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.4 friendliness.png"), plot,
       width = 18, height = 9, units = "cm")

```

Responses to this question were also split, though leaning towards negative, with around `r round(nrow(data %>% filter(friend %in% c("Somewhat friendly to cyclists", "Very cyclist-friendly"))) / nrow(data) * 100)`% saying their suburb was somewhat or very friendly to cyclists, while around `r round(nrow(data %>% filter(friend %in% c("Somewhat unfriendly to cyclists", "Very unfriendly to cyclists"))) / nrow(data) * 100)`% considering their suburb somewhat or very unfriendly.

### Cycling encouragement

```{r}
q <- respondents %>% select(starts_with("3.5")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(encourage = ifelse(.data[[q]] == "No, not all encouraged",
                            "No, not at all encouraged", .data[[q]]),
         encourage = factor(encourage,
                            levels = c("Yes, strongly encouraged", "Yes, to some extent",
                                       "Neutral", "No, not really encouraged",
                                       "No, not at all encouraged")),
         encourage = fct_rev(encourage))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = encourage)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Cycling encouragement",
    y       = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.5 encouragement.png"), plot,
       width = 18, height = 9, units = "cm")

```

Most participants (`r round(nrow(data %>% filter(encourage %in% c("No, not really encouraged", "No, not at all encouraged"))) / nrow(data) * 100)`%) felt that cycling was 'not really' or 'not at all' adequately encouraged and promoted.

### Cycling condition improvements

```{r}
q <- respondents %>% select(starts_with("3.6")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(improv = factor(.data[[q]],
                         levels = c("Much better", "Somewhat better",
                                    "About the same", "Somewhat worse",
                                    "Much worse", "Not sure")))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = improv)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Change in cycling conditions",
    y       = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.6 improvements.png"), plot,
       width = 18, height = 9, units = "cm")

```

Most participants (`r round(nrow(data %>% filter(improv == "About the same")) / nrow(data) * 100)`%) considered that cycling conditions had not changed in their suburb in the past year.

### End-of-trip facilities

```{r}
q <- respondents %>% select(starts_with("3.7")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(endoftrip = factor(.data[[q]],
                            levels = c("Excellent", "Good", "Average", "Poor",
                                       "Very poor", "No end of trip facilities")),
         endoftrip = fct_rev(endoftrip))

```

**`r q`**

```{r fig.width = 8, fig.height = 4}

plot <- ggplot(data, aes(x = endoftrip)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "End of trip facilities",
    y       = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/3.7 end-of-trip.png"), plot,
       width = 18, height = 9, units = "cm")

```

The most common answer (`r round(nrow(data %>% filter(endoftrip == "Average")) / nrow(data) * 100)`%) was 'average', though around `r round(nrow(data %>% filter(endoftrip %in% c("Poor", "Very poor", "No end of trip facilities"))) / nrow(data) * 100)`% answered that end-of-trip facilities were poor, very poor or non-existent.

### Preferred road type


```{r}

col_names <- names(respondents)
q_cols <- col_names[startsWith(col_names, "3.8")]

q <- stringr::str_extract(q_cols[1], ".*\\)")  # part of col name before )

data <- valid_responses %>% 
  select(all_of(q_cols)) %>%
  rename_with(~ str_replace(.x, "^.*?\\):\\s*", "")) %>% 
  pivot_longer(everything(), names_to = "type", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(type = str_replace(type, "paves", "paved"),
         type = type %>% 
           fct_infreq() %>%
           fct_rev())
         

```

**`r q`**

```{r fig.width = 8, fig.height = 5}
plot <- ggplot(data, aes(x = type)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Preferred road type",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/2.8 road type.png"), plot,
       width = 18, height = 12, units = "cm")

```

'Paved off-street multiuse paths' (`r round(nrow(data %>% filter(type == "Paved off-street multiuse paths")) / nrow(data) * 100)`%) and 'paved off-street cycle paths for bikes only' (`r round(nrow(data %>% filter(type == "Paved off-street cycle paths for bikes only")) / nrow(data) * 100)`) were the most-preferred surfaces, indicating a high preference for both separation from traffic and paved surfaces.

### Cycling initiative priorities

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "off-road")          ~ "More off-road paths",
    str_detect(x, "on-road")           ~ "More on-road lanes",
    str_detect(x, "schools")           ~ "Better connections to schools",
    str_detect(x, "shops")             ~ "Better connections to shops",
    str_detect(x, "parks")             ~ "Better connections to parks/pools",
    str_detect(x, "public transport")  ~ "Better connections to PT",
    str_detect(x, "parking")           ~ "More bicycle parking",
    str_detect(x, "separated cycling") ~ "Lack of separation",
    str_detect(x, "speed limits")      ~ "Lower speed limits",
    str_detect(x, "training")          ~ "More training",
    str_detect(x, "signs")             ~ "More signage",
    str_detect(x, "events")            ~ "More events/campaings",
    TRUE                               ~ NA_character_
  )
}

col_names <- names(respondents)
q_cols <- col_names[startsWith(col_names, "3.9")]

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

responses <- data.frame(
  response = str_replace(q_cols, "^.*?:\\s*", ""),  # everything after colon
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- valid_responses %>% 
  select(all_of(q_cols)) %>%
  rename_with(~ str_replace(q_cols, "^.*?:\\s*", "")) %>% 
  pivot_longer(everything(), names_to = "initiative", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(positive = if_else(value %in% c("Very high", "High"), 1L, 0L)) %>%
  mutate(category = categorise(initiative)) %>%
  group_by(category) %>%
  mutate(positive_count = sum(positive, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(category = fct_reorder(category, positive_count),
         category = fct_rev(category),
         value = factor(value,
                        levels = c("Very high",
                                   "High",
                                   "Moderate",
                                   "Low",
                                   "Very low",
                                   "Not a priority")),
         value = fct_rev(value))

caption_text <- responses %>%
  arrange(factor(category, levels = levels(data$category))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

```

**`r q`**

```{r fig.width = 8, fig.height = 5}
plot <- ggplot(data, aes(x = fct_rev(category), fill = value)) +
  geom_bar() +
  scale_fill_brewer(palette = "RdBu") +
  labs(
    x = "Initiative priorities",
    y = "Number of participants",
    fill = "Priority level",
    caption = caption_text
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        plot.caption.position = "plot",
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE))

plot
ggsave(paste0(plot_folder, "/3.9 initiatives.png"), plot,
       width = 18, height = 15, units = "cm")

```

The answer with the highest level of 'High' or 'Very high' priority was 'More off-road paths' (`r round(nrow(data %>% filter(category == "More off-road paths" & value %in% c("Very high", "High"))) / nrow(data %>% filter(category == "More off-road paths")) * 100)`%), indicating the strong desire of many participants for separation from traffic.

### Willingness to detour for dedicated paths

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "Significant")  ~ "Significant",
    str_detect(x, "Considerable") ~ "Considerable",
    str_detect(x, "Moderate")     ~ "Moderate",
    str_detect(x, "Minor")        ~ "Minor",
    str_detect(x, "No detour")    ~ "No detour",
    str_detect(x, "Undecided")    ~ "Undecided",
    TRUE                          ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("3.10")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(category = categorise(.data[[q]]),
         category = factor(category,
                         levels = c("Significant", "Considerable",
                                    "Moderate", "Minor",
                                    "No detour", "Undecided")))

responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  filter(!is.na(category) & !category %in% c("No detour", "Undecided")) %>% 
  arrange(factor(category,
                 levels = c("Significant", "Considerable", "Moderate", 
                            "Minor", "No detour", "Undecided"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")


```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Willingness to detour for dedicated paths",
    y       = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/3.10 detour paths.png"), plot,
       width = 18, height = 9, units = "cm")

```

The most common answer was 'moderate' (`r round(nrow(data %>% filter(category == "Moderate")) / nrow(data) * 100)`%), with around `r round(nrow(data %>% filter(category %in% c("Moderate", "Considerable", "Significant"))) / nrow(data) * 100)`% of participants being willing to make a 'moderate', 'considerable' or 'significant' detour to have access to a dedicated bicycle path.

### Willingness to detour for greenery

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "Significant")  ~ "Significant",
    str_detect(x, "Considerable") ~ "Considerable",
    str_detect(x, "Moderate")     ~ "Moderate",
    str_detect(x, "Minor")        ~ "Minor",
    str_detect(x, "No detour")    ~ "No detour",
    str_detect(x, "Undecided")    ~ "Undecided",
    TRUE                          ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("3.11")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(category = categorise(.data[[q]]),
         category = factor(category,
                         levels = c("Significant", "Considerable",
                                    "Moderate", "Minor",
                                    "No detour", "Undecided")))

responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  filter(!is.na(category) & !category %in% c("No detour", "Undecided")) %>% 
  arrange(factor(category,
                 levels = c("Significant", "Considerable", "Moderate", 
                            "Minor", "No detour", "Undecided"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")


```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Willingness to detour for greenery",
    y       = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/3.11 detour greenery.png"), plot,
       width = 18, height = 9, units = "cm")

```

Participants were less willing to detour for greenery than for separated cycle infrastructure, with only around `r round(nrow(data %>% filter(category %in% c("Moderate", "Considerable", "Significant"))) / nrow(data) * 100)`% being willing to make a 'moderate', 'considerable' or 'significant' detour, and around `r round(nrow(data %>% filter(category == "No detour")) / nrow(data) * 100)`% not being willing to make any detour.

### Willingness to detour for flat route

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "Significant")  ~ "Significant",
    str_detect(x, "Considerable") ~ "Considerable",
    str_detect(x, "Moderate")     ~ "Moderate",
    str_detect(x, "Minor")        ~ "Minor",
    str_detect(x, "No detour")    ~ "No detour",
    str_detect(x, "Undecided")    ~ "Undecided",
    TRUE                          ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("3.12")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(category = categorise(.data[[q]]),
         category = factor(category,
                         levels = c("Significant", "Considerable",
                                    "Moderate", "Minor",
                                    "No detour", "Undecided")))

responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  filter(!is.na(category) & !category %in% c("No detour", "Undecided")) %>% 
  arrange(factor(category,
                 levels = c("Significant", "Considerable", "Moderate", 
                            "Minor", "No detour", "Undecided"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")


```

**`r q`**

```{r fig.width = 8, fig.height = 4}
plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Willingness to detour for flat route",
    y       = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/3.12 detour flat.png"), plot,
       width = 18, height = 9, units = "cm")

```

Results for this question were varied, with the most common answer being 'moderate' (`r round(nrow(data %>% filter(category == "Moderate")) / nrow(data) * 100)`%), and around `r round(nrow(data %>% filter(category %in% c("Moderate", "Considerable", "Significant"))) / nrow(data) * 100)`% of participants being willing to make a 'moderate', 'considerable' or 'significant' detour to achieve a flatter trip, but also around `r round(nrow(data %>% filter(category == "No detour")) / nrow(data) * 100)`% not being willing to make any detour to avoid hills.
