---
title: "Cycling survey results"
author: "Steve"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(fs)
library(readxl)
library(scales)

plot_folder <- "../Bendigo survey/plots"
if (!dir.exists(plot_folder)) dir_create(plot_folder)

```

```{r load_data, message=FALSE}
surveyFile <- "../Bendigo survey/responses-8gi6nyx69dt6-2025-07-10T04_40_41.798Z.xlsx"

sheet_names <- excel_sheets(surveyFile)

spot_sheet_names <- c("Most stressful intersection(s) ",
                      "My favorite spot(s) along the r",
                      "My least favorite spot(s) along")

route_sheet_names <- sheet_names[!sheet_names %in% c("Respondents", 
                                                     spot_sheet_names)]

respondents <- read_excel(surveyFile, sheet = "Respondents")

routes <- lapply(route_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$destination <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()

spots <-  lapply(spot_sheet_names, function(sheet) {
  data <- read_excel(surveyFile, sheet = sheet)
  data$type <- sheet  # add sheet name as a new column
  return(data)
}) %>%
  bind_rows()


```

# Introduction
There were `r nrow(respondents)` survey responses, though some were incomplete.  

The survey is divided into 3 parts:
- sociodemographic characteristics
- cycling experiences
- cycling preferences
 
# 1. Sociodemographic characteristics
This section provides general information about the survey participants.

## Ridership

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "once a week")  ~ "Weekly",
    str_detect(x, "once a month") ~ "Monthly",
    str_detect(x, "occasionally") ~ "Occasionally",
    str_detect(x, "not cycled")   ~ "Not cycled",
    is.na(x)                      ~ "Did not answer",
    TRUE                          ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("Are you a bicycle rider")) %>% names()

data <- respondents %>%
  select(all_of(q)) %>%
  mutate(category = categorise(.data[[q]])) %>%
  mutate(category = factor(category,
                           levels = c("Weekly", "Monthly", "Occasionally",
                                      "Not cycled", "Did not answer")))
responses <- respondents %>%
  distinct(response = .data[[q]]) %>%
  mutate(category = categorise(response))

caption_text <- responses %>%
  filter(category != "Did not answer") %>% 
  arrange(factor(category,
                 levels = c("Weekly", "Monthly", "Occasionally", "Not cycled"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

valid_responses <- respondents %>%
  mutate(category = categorise(.data[[q]])) %>%
  filter(category %in% c("Weekly", "Monthly", "Occasionally"))
```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = category)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Category",
    y       = "Number of respondents",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.0 ridership.png"), plot,
       width = 18, height = 9, units = "cm")

```

Out of `r nrow(respondents)` survey participants, `r nrow(respondents) - nrow(valid_responses)` answered that they had not cycled in the past year, or did not answer. These were excluded from further survey questions.

The results presented below for the remainder of the sociodemographic questions exclude participants who did not answer the relevant question.


### Gender

```{r}
q <- respondents %>% select(starts_with("1.1")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(gender = factor(.data[[q]],
                         levels = c("Woman/Female", "Man/Male", 
                                    "Other", "Prefer not to say")))

```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = gender)) +
  geom_bar(fill = "steelblue") +
  labs(
    x       = "Gender",
    y       = "Number of participants"
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.1 gender.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Age

```{r, warning=FALSE}
q <- respondents %>% select(starts_with("1.2")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  
  # anomalous answers - assign midpoint
  mutate(age_num = case_when(
    str_detect(.data[[q]], "Mid 40s") ~ 45,
    str_detect(.data[[q]], "50’s") ~ 55,
    TRUE ~ as.numeric(.data[[q]])
  )) %>%
  # ambiguous answers/errors (60+, 565) - exclude
  filter(!is.na(age_num) & age_num <= 100) %>%
  
  mutate(age_group = cut(
  age_num,
      breaks = c(17, 20, 30, 40, 50, 60, 70, 80, Inf),
      labels = c("18-20", "21-30", "31-40", "41-50",
                 "51-60", "61-70", "71-80", "80+"),
      right = TRUE,   # intervals are closed on the right
      include.lowest = TRUE
    ))
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = age_group)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Age group",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.2 age.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Living arrangements

```{r}
q <- respondents %>% select(starts_with("1.3")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(living_arrangement = .data[[q]] %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev() %>%   # reverse so largest is on top in horizontal bar chart
           fct_relevel("Prefer not to say", after = 0)# push to first (bottom in horizontal bars)
  )
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = living_arrangement)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Living arrangement",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.3 living.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Cars

```{r}
q <- respondents %>% select(starts_with("1.4")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_cars = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = num_cars)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of cars",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.4 cars.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Bicycles

```{r}
q <- respondents %>% select(starts_with("1.5")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(num_bikes = factor(.data[[q]], 
                           levels = sort(unique(.data[[q]])))) 
    
```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = num_bikes)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Number of bicycles",
    y = "Number of participants"
  ) +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.5 bikes.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Educational qualifications

```{r}
categorise <- function(x, y) {
  case_when(
    str_detect(x, "Year 10")         ~ "Year 10 or less",
    x == "Other" & (str_detect(y, "11") | str_detect(y, "12")) ~ "Year 11 or 12",
    str_detect(x, "Diploma") ~ "Diploma/certificate",
    str_detect(x, "Undergraduate")   ~ "Undergraduate",
    str_detect(x, "Postgraduate")    ~ "Postgraduate",
    x == "Other"                     ~ "Other",
    TRUE                             ~ NA_character_
  )
}

q <- respondents %>% select(starts_with("1.6")) %>% names()
col_names <- names(respondents)
next_col <- col_names[which(col_names == q) + 1]

data <- valid_responses %>%
  select(all_of(c(q, next_col))) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(qualif = categorise(.data[[q]], .data[[next_col]])) %>%
  mutate(qualif = factor(qualif,
                           levels = c("Year 10 or less", "Year 11 or 12", "Diploma/certificate",
                                      "Undergraduate", "Postgraduate", "Other")))

responses <- data %>%
  distinct() %>%
  rename(response = 1) %>%
  arrange(qualif) %>%
  filter(str_detect(qualif, "Diploma") | str_detect(qualif, "Undergraduate") | 
           str_detect(qualif, "Postgraduate"))

caption_text <- responses %>%
  arrange(factor(qualif,
                 levels = c("Diploma/certificate", "Undergraduate", "Postgraduate"))) %>%
  mutate(full = paste0(qualif, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

```

**`r q`**

```{r}
plot <- ggplot(data, aes(x = qualif)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Educational qualification",
    y = "Number of participants",
    caption = caption_text
  ) +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

plot
ggsave(paste0(plot_folder, "/1.6 qualification.png"), plot,
       width = 18, height = 9, units = "cm")

```

### Occupation

```{r}
q <- respondents %>% select(starts_with("1.7")) %>% names()

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]]) & !(.data[[q]] == "-")) %>%
  mutate(occup = tolower(.data[[q]])) %>%
  mutate(occup = case_when(
    str_detect(occup, "retired") & !str_detect(occup, "semi") ~ "retired",
    occup == "admin" ~ "administration",
    occup == "physio" ~ "physiotherapist",
    TRUE             ~ occup
  )) %>%
  mutate(occup = str_to_sentence(occup)) %>%
  mutate(occup = case_when(
    occup == "Ceo"                           ~ "CEO",
    occup == "Council officer (not cogb)"    ~ "Council officer (not CoGB)",
    occup == "Early support coordinator - transport accident commission" ~
      "Early support coordinator - Transport Accident Commission",
    occup == "Gis officer"                   ~ "GIS officer",
    occup == "It administrator/ gis analyst" ~ "IT administrator / GIS analyst",
    occup == "Gp"                            ~ "GP",
    occup == "Mri applications specialist"   ~ "MRI applications specialilst",
    occup == "Public servant dtp"            ~ "Public servant DTP",
    TRUE                                     ~ occup
  )) %>%
  group_by(occup) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  arrange(desc(n), occup)

data_plot <- data %>%
  filter(n >= 3) %>%
  mutate(occup = occup %>%
           fct_infreq() %>%  # order by frequency (largest first)
           fct_rev())  # reverse so largest is on top in horizontal bar chart

others_text <- data %>%
  filter(n < 3) %>%
  mutate(text = paste0(occup, " (", n, ")")) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = ", ")


```

**`r q`**

```{r}
plot <- ggplot(data_plot, aes(x = occup)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Occupation",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.7 occupation.png"), plot,
       width = 18, height = 9, units = "cm")

```

Other occupations not included in the plot above: 
`r others_text`.

### Income

```{r}
q <- respondents %>% select(starts_with("1.8")) %>% names()

format_income <- function(x) {
  str_replace_all(x, 
                  "(\\d+)",  # match any sequence of digits
                  function(y) paste0("$", comma(as.numeric(y))))
}

data <- valid_responses %>%
  select(all_of(q)) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(income_label = format_income(.data[[q]])) %>%
  mutate(income_order = case_when(
    str_detect(income_label, "Below") ~ 0,
    str_detect(income_label, "Above") ~ 200000,  # arbitrarily above highest range
    str_detect(income_label, "I prefer") ~ Inf,
    TRUE ~ as.numeric(str_extract(income_label, "\\d+"))
  )) %>%
  mutate(income_label = fct_reorder(income_label, income_order) %>%
           fct_rev())
  
```

**`r q`**

```{r}

plot <- ggplot(data, aes(x = income_label)) +
  geom_bar(fill = "steelblue") +
  labs(
    x = "Household income",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw()

plot
ggsave(paste0(plot_folder, "/1.8 income.png"), plot,
       width = 18, height = 9, units = "cm")

```

# 2. Cycling experiences

This section provides information about routes drawn by participants.

### Route purpose
Participants were asked to describe the purpose of their route.

```{r}
clean_destination <- function(x) {
  x <- case_when(
    x == "To public transport stop-statio" ~ "To public transport stop/station",
    TRUE ~ x
  )
  x <- gsub("-", "/", x)
  x <- gsub(", ", "/", x)
  x <- factor(
    x,
    levels = c(
      "To work/commute", "To college/university/classes",
      "To shops/tasks/errands", "To visit family/friends",
      "To public transport stop/station", "To other destinations"
    )
  )
  return(x)
}

data <- routes %>% 
  dplyr::select(`Respondent ID`, destination, `Please tell us the type of destination:`) %>%
  mutate(destination = clean_destination(destination) %>%
           fct_rev())

others_text <- data %>%
  filter(!is.na(`Please tell us the type of destination:`)) %>%
  group_by(`Please tell us the type of destination:`) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(`Please tell us the type of destination:`, " (", n, ")"),
                       `Please tell us the type of destination:`)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")


plot <- ggplot(data, aes(x = destination, fill = destination)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Trip purpose",
    y = "Number of participants"
  ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none")

plot
ggsave(paste0(plot_folder, "/2.0 purpose.png"), plot,
       width = 18, height = 9, units = "cm")

```

Other destinations given:<br>
`r others_text`


### Reasons for cycling

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "exercise")   ~ "Exercise",
    str_detect(x, "enjoy")      ~ "Enjoyment",
    str_detect(x, "available")  ~ "Availability",
    str_detect(x, "convenient") ~ "Convenience",
    str_detect(x, "peers")      ~ "Peers",
    str_detect(x, "Other")      ~ "Other",
    TRUE                        ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "1.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

responses <- data.frame(
  response = str_replace(q_cols, "^.*?:\\s*", ""),  # everything after colon
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Exercise", "Enjoyment", "Availability",
                                            "Convenience", "Peers", "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Exercise", "Enjoyment", "Availability",
                                            "Convenience", "Peers", "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(category = fct_rev(category),
                        destination = fct_rev(destination)), 
               aes(x = category, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Reason for cycling",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.1 reason.png"), plot,
       width = 18, height = 12, units = "cm")


```

Reasons given for 'Other':<br>
`r others_text`

### Reasons for selecting route

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "safe")          ~ "Safe",
    str_detect(x, "greenery")      ~ "Green",
    str_detect(x, "flat")          ~ "Flat",
    str_detect(x, "traffic")       ~ "Traffic",
    str_detect(x, "intersections") ~ "Intersections",
    str_detect(x, "speed")         ~ "Speed limits",
    str_detect(x, "connectivity")  ~ "Connectivity",
    str_detect(x, "shortest")      ~ "Short",
    str_detect(x, "cyclists")      ~ "Other cyclists",
    str_detect(x, "designated")    ~ "Designated",
    str_detect(x, "Other")         ~ "Other",
    TRUE                           ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "2.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], ".*\\)")  # part of col name before )

responses <- data.frame(
  response = str_replace(q_cols, "^.*?\\):\\s*", ""),  # everything after ):
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?\\):\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Safe", "Green", "Flat", "Traffic",
                                            "Intersections", "Peers", "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Safe", "Green", "Flat", "Traffic", "Intersections",
                            "Speed limits", "Connectivity", "Short", 
                            "Other cyclists", "Designated", "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.1.`r q`**

```{r fig.height = 12}
plot <- ggplot(data %>%
                 mutate(category = fct_rev(category),
                        destination = fct_rev(destination)),
               aes(x = category, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Reason for choosing route",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.2 route choice.png"), plot,
       width = 18, height = 18, units = "cm")


# # facet option (if prefered - would need 'fig.height = 12' added to {r} above)
# # could roll this out to all relevant parts of section 2
# plot <- ggplot(data %>% mutate(category = fct_rev(category),
#                                destination = fct_rev(destination)),
#                aes(x = category)) +
#   geom_bar(aes(fill = destination)) +   # still map fill to destination
#   facet_wrap(~ fct_rev(destination), ncol = 1, scales = "free_y") +
#   scale_fill_brewer(palette = "Set1", direction = -1) +  # 6 colors
#   coord_flip() +
#   labs(
#     x = "Reason for choosing route",
#     y = "Number of participants",
#     fill = "Trip purpose",
#     caption = caption_text
#   ) +
#   theme_bw() +
#   theme(plot.caption = element_text(hjust = 0),
#         legend.position = "none")
# 
# plot
# ggsave(paste0(plot_folder, "/2.1.2 route choice.png"), plot,
#        width = 18, height = 36, units = "cm")
```

Reasons given for 'Other':<br>
`r others_text`.

### Maximum distance

```{r}
q <- routes %>% select(starts_with("3.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(distance = str_replace(.data[[q]], "(km)", "\\1\n"),  # insert line break after 'km'
         distance = factor(distance,
                         levels = c("Up to 5 km\n (approximately 15 minutes)",
                                    "Up to 10 km\n (approximately 30 minutes)",
                                    "Up to 20 km\n (approximately 1 hour)",
                                    "Up to 30 km\n (approximately 1.5 hour)",
                                    "More than 30 km\n (more than 1.5 hour)")),
         
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(distance = fct_rev(distance),
                        destination = fct_rev(destination)), 
               aes(x = distance, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Distance willing to travel",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.3 distance.png"), plot,
       width = 18, height = 12, units = "cm")

```

### Unfavourable route aspects

```{r}
categorise <- function(x) {
  case_when(
    str_detect(x, "unsafe")        ~ "Unsafe",
    str_detect(x, "separation")    ~ "Not separated",
    str_detect(x, "greenery")      ~ "Not green",
    str_detect(x, "traffic")       ~ "Traffic",
    str_detect(x, "intersections") ~ "Intersections",
    str_detect(x, "speed")         ~ "Speeds",
    str_detect(x, "connectivity")  ~ "Poor connectivity",
    str_detect(x, "surface")       ~ "Surface",
    str_detect(x, "steep")         ~ "Steep",
    str_detect(x, "Other")         ~ "Other",
    TRUE                           ~ NA_character_
  )
}

col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "4.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]  # 'Please specify' column
other_col2 <- str_replace(other_col, "\\d+$", 
                          function(n) as.character(as.integer(n) + 1)) # 'Please specify' +1

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

responses <- data.frame(
  response = str_replace(q_cols, "^.*?:\\s*", ""),  # everything after colon
  stringsAsFactors = FALSE
) %>%
  mutate(category = categorise(response))

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "reason", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    category = categorise(reason),
    category = factor(category, levels = c("Unsafe", "Not separated", "Not green",
                                           "Traffic", "Intersections", "Speeds",
                                           "Poor connectivity", "Surface", "Steep",
                                           "Other"))
  )

caption_text <- responses %>%
  filter(response != "Other") %>%
  arrange(factor(category,
                 levels = c("Unsafe", "Not separated", "Not green",
                            "Traffic", "Intersections", "Speeds",
                            "Poor connectivity", "Surface", "Steep",
                            "Other"))) %>%
  mutate(full = paste0(category, ": ", response)) %>%
  pull(full) %>%
  paste(collapse = "\n")

others_text <- routes %>%
  select(all_of(c(other_col, other_col2))) %>%
  pivot_longer(everything(), values_to = "other") %>%
  select(other) %>%
  filter(!is.na(other)) %>%
  group_by(other) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(text = ifelse(n > 1,
                       paste0(other, " (", n, ")"),
                       other)) %>%
  distinct(text) %>%
  arrange(text) %>%
  pull(text) %>%
  paste(collapse = "<br>")

```

**2.1.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(category = fct_rev(category),
                        destination = fct_rev(destination)), 
               aes(x = category, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Unfavourable route aspects",
    y = "Number of participants",
    fill = "Trip purpose",
    caption = caption_text
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.4 unfavourable.png"), plot,
       width = 18, height = 18, units = "cm")

```

Other unfavourable route aspects given:<br>
`r others_text`

### Frequency of cycling route

```{r}
q <- routes %>% select(starts_with("5.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(frequency = str_replace(.data[[q]], "-", "-\n"),  # insert line break after first dash
         frequency = factor(frequency,
                         levels = c("Frequently-\n 5 days or more per week",
                                    "Regularly-\n at least 4 days per week",
                                    "Often-\n 1-3 days per week",
                                    "From time to time-\n 1-4 days per month")),
         
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(frequency = fct_rev(frequency),
                        destination = fct_rev(destination)), 
               aes(x = frequency, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Frequency of cycling route",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.5 frequency.png"), plot,
       width = 18, height = 12, units = "cm")

```

### Satisfaction level

```{r}
q <- routes %>% select(starts_with("6.")) %>% names()

data <- routes %>%
  select(all_of(q), destination) %>%
  filter(!is.na(.data[[q]])) %>%
  mutate(satisf = str_replace(.data[[q]], "-", "-\n"),  # insert line break after first dash
         satisf = factor(satisf,
                         levels = c("Very satisfied",
                                    "Satisfied",
                                    "Neutral",
                                    "Dissatisfied",
                                    "Very dissatisfied")),
         destination = clean_destination(destination))

```

**2.1.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(satisf = fct_rev(satisf),
                        destination = fct_rev(destination)), 
               aes(x = satisf, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Satisfaction with route",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.6 satisfaction.png"), plot,
       width = 18, height = 12, units = "cm")

```


### Accompanying people

```{r}
col_names <- names(routes)
q_cols <- col_names[startsWith(col_names, "7.")]
last_idx <- which(col_names == tail(q_cols, 1))
other_col <- col_names[last_idx + 1]

q <- stringr::str_extract(q_cols[1], "^[^:]+")  # part of col name before colon

data <- routes %>% 
  select(all_of(q_cols), destination) %>%
  rename_with(~ str_replace(.x, "^.*?:\\s*", "")) %>% 
  pivot_longer(cols = -destination, names_to = "people", values_to = "value") %>%
  filter(value == TRUE) %>%
  mutate(
    destination = clean_destination(destination),
    people = factor(people, levels = c("My friend(s)", "My partner", "My kids",
                                            "Other family members", "I cycle alone"))
  )


```

**2.`r q`**

```{r}
plot <- ggplot(data %>%
                 mutate(people = fct_rev(people),
                        destination = fct_rev(destination)), 
               aes(x = people, fill = destination)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(
    x = "Accompanying people",
    y = "Number of participants",
    fill = "Trip purpose"
  ) +
  coord_flip() +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0), 
        legend.position = "bottom") +
  guides(fill = guide_legend(reverse = TRUE, nrow = 3))

plot
ggsave(paste0(plot_folder, "/2.1.7 accompanying.png"), plot,
       width = 18, height = 12, units = "cm")


```

```{r}
spot_numbers <- spots %>%
  mutate(category = case_when(
    str_detect(type, "My favorite")    ~ "Favourite",
    str_detect(type, "least favorite") ~ "Least favourite",
    str_detect(type, "stressful")       ~ "Stressful",
    TRUE                                ~ NA_character_
  )) %>%
  group_by(category) %>%
  summarise(n = n()) %>%
  ungroup()

fav_no <- spot_numbers[spot_numbers$category == "Favourite", "n"]

fav_no <- spot_numbers %>% filter(category == "Favourite") %>% pull(n)
```


### Favourite spot

**2.2 My favorite spot(s) along the route**

A total of `r spot_numbers %>% filter(category == "Favourite") %>% pull(n)` favourite spots were marked by participants.

### Stressful intersections

**2.3 Most stressful instersection(s)**

A total of `r spot_numbers %>% filter(category == "Stressful") %>% pull(n)` stressful intersections were marked by participants.

### Least favourite spot

**2.4 My least favorite spot(s) along the route**

A total of `r spot_numbers %>% filter(category == "Least favourite") %>% pull(n)` least favourite spots were marked by participants.

### Barriers and challenges

```{r}
### UP TO HERE

# To do section 2.5 from 'routes'.  It's a multi-choice one with multiple options; can base it on 
# 2.1.2 (reason for selecting routes).  Will need a caption, and 'other' text

```

